---
title: "PML_Project"
output: html_document
---

Starting with the analysis, we load de library we will need.

```{r}
library(caret)
library(ggplot2)
library(rattle)
```

Load de data en do a summary to understand the type of data. I fix the empty data to NA.

```{r}
data <- read.csv("pml-training.CSV",na.strings=c("NA",""))
summary(data)
```

We make some graphics to find the relations.

```{r}
qplot(pitch_forearm,roll_forearm,data=data,colour=classe)
qplot(pitch_forearm,roll_belt,data=data,colour=classe)
```

Indentify the variables wich I won't use, the firt 7 variables and the variables with NA.
```{r}
col <- c("X","user_name","cvtd_timestamp","raw_timestamp_part_1","raw_timestamp_part_2","new_window","num_window")


for(i in 1:dim(data)[2]){
  if(sum(is.na(data[,i])) > 0){
    col <- c(col,names(data[i]))
  }
}

datacln <- data[,!(names(data) %in% col)]

```

Split the data in training and testing, using the  createFolds function I generated 3 sets of traning-testing.

```{r}
inTrain <- createDataPartition(y=datacln$classe, p=0.6,list=FALSE)

set.seed(200)
trfs <- createFolds(y=datacln$classe,k=3,list=TRUE,returnTrain=TRUE)

training1 <- datacln[trfs[[1]],]
training2 <- datacln[trfs[[2]],]
training3 <- datacln[trfs[[3]],]
testing1 <- datacln[-trfs[[1]],]
testing2 <- datacln[-trfs[[2]],]
testing3 <- datacln[-trfs[[3]],]
```

The first model is simple with rpart method.
```{r}
set.seed(300)
fit1 <- train(classe~.,method="rpart", data=training1)
predict1 <- predict(fit1, newdata=testing1)
confusionMatrix(predict1,testing1$classe)
```

I create another model rpart with a cross valitation train control.
```{r}
set.seed(300)
fit2 <- train(classe~.,method="rpart", data=training2,trControl=trainControl(method="cv",number=4))
predict2 <- predict(fit2, newdata=testing2)
confusionMatrix(predict2,testing2$classe)
```

Finally, another model with rando forest.
```{r}
set.seed(300)
fit3 <- train(classe~.,method="rf", data=training3,trControl=trainControl(method="cv",number=4), preProcess=c("center", "scale"))
predict3 <- predict(fit3, newdata=testing3)
confusionMatrix(predict3,testing3$classe)

```

The randon forest is way better than the others according to the accuracy, is the final model.


Test the model in the test data.

```{r}
pmltest <- read.csv("pml-testing.CSV",na.strings=c("NA",""))
predict_test <- predict(fit3, newdata=pmltest)
predict_test

```

